---
title: "Final Paper"
author: "Liz Gorrill"
date: "11/28/2021"
output: pdf_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```


# Abstract

The Surfrider Foundation, Massachusetts Chapter is a non-profit organization that "is dedicated to the protection and enjoyment of the world's ocean, waves and beaches, for all people, through a powerful activist network" (Surfrider Foundation, 2021). Each year beach cleanups serve as the primary form of volunteer work and the events provide a space to collect data through citizen science as well as discuss the impacts of the trash which is collected. The coronavirus (COVID-19) pandemic posed an obstacle to beach cleanups in 2021 with gatherings being against local regulations until June 2021 and volunteer comfort with gatherings being hindered through the entirety of 2021. In order to continue making progress on beach cleanups and citizen science data collection of beach trash, the Connecting Our Coastline (COC) do-it-yourself (DIY) beach cleanup initiative was introduced with the goal of at least one beach cleanup in each coastal community of Massachusetts. This paper serves as an assessment of the year long effort of COC and the implications of the results that were found through the data collected at each DIY beach cleanup.

# Introduction

Discuss the Surfrider Foundation, its mission and the MA chapter in specific, ie. types of events and community involvement

Discuss the types of trash most commonly found on Massachusetts beaches and why cataloging the types of trash that are found is important - accountability, marine entanglements and starvation, 

In an effort to maintain volunteer involvement and continue the commitment to keeping Massachusetts beaches clean, the Surfrider Foundation, Massachusetts Chapter executive committee created DIY beach cleanup materials in November 2020 to encourage chapter volunteers and residents of coastal communities to get outside and 

# Methods

In November 2020, the executive commitee of the Surfrider Foundation, Massachusetts Chapter, put together materials including data collection cards, beach cleanup tips and disclaimers in the DIY page of the chapters website in order to provide people with a social distancing option for keeping Massachusetts beaches clean. To gain involvement in the DIY cleanups, an initiative was held in December 2020 where people who signed up, completed a DIY beach cleanup and returned their data cards were rewarded with a gift from the chapter. This successful initiative, was then followed by a period of limited chapter related beach cleanups from January 2021 through March 2021 due to increased restrictions surrounding the COVID-19 pandemic in Massachusetts as well as restrictions put in place by Surfrider Foundation, National.

The Connecting our Coastlines (COC) initiative began in April 2021, with the initiative being announced during the monthly chapter meeting on April 7, 2021. Along with the announcement the materials previously published in November 2020 were shared to chapter volunteers through the chapter e-mail and were made readily available to the public through social media postings on Facebook and Instagram. On June 8, 2021 as part of World Oceans Day, the chapter held the first in person beach cleanup event (since the inception of the COVID-19 pandemic in Massachusetts in March 2020) where volunteers were asked to sign up for time slots to allow for social distancing and time for disinfection of buckets and clipboards between groups. This cleanup served as a way to gain more interest in COC and was the beginning of six months of consistent DIY beach cleanups along with ten (10) chapter beach cleanups. 

Those participating in DIY cleanups and chapter cleanups were asked to collect data regarding the trash that they collected, cataloging the type of trash and the amount of each type collected (individual pieces). Once the DIY or chapter beach cleanups were completed, the collected data was submitted either by the individual through a google form which allowed them to input all of the data, or submitted via mail or e-mail to the chapter executive committee for data entry. The data entry process included the entry of trash types, trash amounts, total weight of cleanup, cleanup location (beach, latitude, longitude and town) and providing follow-up volunteer emails if requested on the submitted data sheet. From the online database, sensitive volunteer information was removed from the downloaded working spreadsheet for the volunteers security. 

#Discussion

Discuss cleanup locations
Discuss Trash totals
Discuss the consequences of types of trash that were found - particularly those which were commonly found
Discuss the current work Surfrider is doing to address the types of trash seen in Massachusetts


```{r}

library(maps)
library(ggmap)
MA <- map_data("state", "Massachusetts")

COC <- read.csv("MA_Cleanup_Data.csv", stringsAsFactors = TRUE)

ggplot()+geom_polygon(data = MA, aes(x=long, y=lat, group = group), fill = "grey", alpha = 0.6) +
    geom_point(data = COC, aes(x=long, y = lat, size = Weight_lbs, color = Weight_lbs)) + coord_map() + xlab("Longitude") + ylab("Latitude") 
  


 


```

```{r}
library(dplyr)
library(plotly)

COC2 <- COC %>%
  arrange(Weight_lbs) %>%
  mutate(name=factor(Town, unique(Town))) %>%
  mutate(mytext = paste(
    "Town: ", name, "\n",
    "Pounds of Trash Removed: ", Weight_lbs, sep = ""))
  

coc_map <- COC2 %>%
  ggplot() + geom_polygon(data = MA, aes(x = long, y = lat, group = group), fill = "grey", alpha = 0.6) +
  geom_point(aes(x=long, y = lat, size = Weight_lbs, color = Weight_lbs, text = mytext, alpha = 2)) +
  scale_size_continuous(range = c(1,10)) +
  theme_void() +
  coord_map() + 
  labs(title = "Connecting our Coastlines", subtitle = "Surfrider Foundation Massachusetts Chapter beach cleanup data ", color = "Weight (pounds)")
 


coc_map <- ggplotly(coc_map, tooltip = "text")

coc_map

api_create(coc_map, filename = "COC_Locations")


```

```{r}
library(tidyr)

COC_long <- COC%>%
  pivot_longer(cols = c(starts_with("Cig"), starts_with("Plastic"), starts_with("Foam"), starts_with("Ball"), starts_with("Fish"), starts_with("Hyg"), starts_with("Paper"), starts_with("Treated"), starts_with("Wood"), starts_with("Metal"), starts_with("Glass"),  starts_with("Tire"), starts_with("Fire"), starts_with("Tarp"),  starts_with("Shot")),
               names_to = "Trash_Type", values_to = "Trash_Amount") 

ggplot(COC_long, aes(x = Trash_Type, y = Trash_Amount)) + 
    geom_segment( aes(x=Trash_Type, xend=Trash_Type, y=0, yend=Trash_Amount), color="skyblue") +
  geom_point( color="blue", size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  )

###sort data by type, create bar chart of total trash
```



```{r}
   
  
COC_Sum <-   aggregate(COC_long$Trash_Amount, by=list(Trash_Type=COC_long$Trash_Type), FUN=sum, na.rm = TRUE)

coc_bc <- plot_ly(COC_Sum, x = ~Trash_Type, y = ~x, type = "bar", name = "Trash Type Distribution") %>%
    layout(title = 'Beach Trash Totals 2021', xaxis = list(title = 'Trash Type'), 
         yaxis = list(title = 'Trash Amount (# of Individual Pieces)'))
coc_bc

api_create(coc_bc, filename = "COC_Trash_Totals")

```


```{r}

ggplot(COC_Sum, aes(x = Trash_Type, y = x)) + 
    geom_segment( aes(x=Trash_Type, xend=Trash_Type, y=0, yend=x), color="skyblue") +
  geom_point( color="blue", size=4, alpha=0.6) +
  theme_light() +
  coord_flip() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.border = element_blank(),
    axis.ticks.y = element_blank()
  ) +
  xlab("Trash Type") + ylab("Amount (Individual Pieces)")


```

```{r}

ggplot(COC_long, aes(fill = Trash_Type, y = Trash_Amount, x = Beach)) +
  geom_bar(position = "stack", stat = "identity") + 
  theme(axis.text.x = element_text(size = 5, angle = 45, hjust = 1), legend.key.size = unit(.25, "cm")) +
  ylab("Trash Amount (Individual Pieces)") 


```
